---
title: "STAT 107 Final Project"
author: "Ash Mansour, Jenny Zhang, Adam Rui, Richard Zhou"
format: html
editor: visual
---

# Introduction

## Objective

\[Describe purpose of analysis and the dataset\]

## Dataset Overview

-   Source: \[Source\]

-   Description: \[Key attributes of dataset\]

-   Variables: \[Describe variables/columns\]

# Data Preparation

## Cleaning the Data

```{r}
# load and clean data
```
data <- load.csv(career_change_prediction_dataset.csv
\[Quick summary of cleaning\]

## Summary Statistics

```{r}
# summary()
```

\[Maybe a quick summary if anything interesting? if not, not important\]

# Data Visualization

## Analysis #1 (REPLACE WITH APPROPRIATE NAME)

```{r}
# do the thingy
```

\[Summary of analysis\]

## Analysis #2 (REPLACE WITH APPROPRIATE NAME)

```{r} # do the thingy}
```

\[Summary of analysis\]

## Analysis #3 (REPLACE WITH APPROPRIATE NAME)

```{r} # do the thingy}
```

\[Summary of analysis\]

# Key Findings and Summary

## Findings:

-   \[Quick summary of findings from analysis #1\]

-   \[Quick summary of findings from analysis #2\]

-   \[Quick summary of findings from analysis #3\]

## Summary:

\[Summary of entire project\]
