---
title: "STAT 107 Final Project"
author: "Ash Mansour, Jenny Zhang, Adam Rui, Richard Zhou"
format: html
editor: visual
---

# Introduction

## Objective

\[Describe purpose of analysis and the dataset\]

## Dataset Overview

### Description (Source: [kaggle.com](https://www.kaggle.com/datasets/jahnavipaliwal/field-of-study-vs-occupation)):

This dataset of over 30,000 entries of 22 attributes regarding each individual's personal details, job satisfaction, and more, provides excellent insight into the reasons individuals change careers. The dataset focuses on the possible sources of influence that could push someone to both change careers, and also to stay in one.

# Data Preparation

## Cleaning the Data

```{r}
data <- read.csv("data/career.csv")

data <- data[, c(1, 2, 3, 4, 5, 6, 7, 8, 9, 11, 12, 13, 15, 16, 21, 23)]

colnames(data) <- c("field_of_study",              # categorical string: field of study during education
                    "current_occupation",          # categorical string: current industry employed in
                    "age",                         # integer: current age
                    "gender",                      # categorical string: gender of individual
                    "years_experience",            # integer: # of years having been in the workforce
                    "education_level",             # ordinal string: highest level of education completed
                    "industry_growth_rate",        # ordinal string: growth rate of the current industry employed in
                    "job_satisfaction",            # integer (1-10 scale): rating of current job satisfaction
                    "work_life_balance",           # integer (1-10 scale): rating of perceived work-life balance
                    "salary",                      # integer: annual salary in USD
                    "job_security",                # integer (1-10 scale): rating of perceived job security
                    "career_change_interest",      # boolean (1 = Yes, 0 = No): whether interested in changing industries
                    "family_influence",            # ordinal string: degree of influence family had on career choice
                    "mentor_availability",         # boolean (1 = Yes, 0 = No): access to a mentor in current occupation
                    "career_changes",              # integer: # of career changes in the past
                    "likely_to_change_occupation") # boolean (1 = Yes, 0 = No): whether likely to change occupation
```

Although many attributes are very useful, given that this analysis won't require all 22 of them, there's no need to include each and every one. We also changed each variable name to better fit common naming conventions, and to better clarify their purpose.

## Summary Statistics

```{r}
# summary(data)
```

\[Maybe a quick summary if anything interesting? if not, not important\]

# Data Visualization

## Analysis #1 (REPLACE WITH APPROPRIATE NAME)

```{r}
# do the thingy
```

\[Summary of analysis\]

## Analysis #2 (REPLACE WITH APPROPRIATE NAME)

```{r}
# do the thingy
```

\[Summary of analysis\]

## 

# Analysis #3 - Multiple Logistic Regression Analysis - Odds ratio of career change interest

```{r}
model1 <- glm(Career.Change.Interest ~  
                 Gender+Family.Influence
                 +Current.Occupation+Job.Security,
                    family = binomial, dd)
coef <- summary(model1)$coef[-1,]
CI <- round(exp(confint(model1))[-1,],2)
OR <- round(exp(coef[,1]),2)
p.value = round(coef[,4],4)
ORmtx <- data.frame(OR, Lower = CI[,1], Upper=CI[,2],
                    p.value ) 
ORmtx$Index <- seq(1:nrow(ORmtx))

ORmtx$Label <- c("Gender:Male vs. Female",
                 "Family Influence: Low vs. None",
                 "Family Influence: Medium vs. None",
                 "Family Influence: High vs. None",
                 "Current Occupation:Artist vs. Software Developer",
                 "       Mechanical Engineer vs. Software Developer",
                 "          Business Analyst vs. Software Developer",
                 "                    Lawyer vs. Software Developer",
                 "                 Economist vs. Software Developer",
                 "                   Teacher vs. Software Developer",
                 "                    Doctor vs. Software Developer",
                 "                  Biologist vs. Software Developer",
                 "Job.Security"
                 )
ORmtx[,c("OR","Lower","Upper","p.value" )]
```

```{r}
ggplot(ORmtx, aes(y = Index, x = OR)) +
  geom_point(shape = 18, size = 5) +  
  geom_errorbarh(aes(xmin = Lower, xmax = Upper), height = 0.25) +
  geom_vline(xintercept = 1, color = "red", linetype = "dashed",  alpha = 0.5) +
  scale_y_continuous(name = "", breaks=1:13, labels = ORmtx$Label, trans = "reverse") +
  xlab("Odds Ratio (95% CI)") + 
  ylab(" ") + 
  theme_bw() +
  theme(panel.border = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"),
        axis.text.y = element_text(size = 12, colour = "black"),
        axis.text.x.bottom = element_text(size = 12, colour = "black"),
        axis.title.x = element_text(size = 12, colour = "black"))

```

\[Summary of analysis\]

The multiple logistic regression shows that:

1.  **Gender**: no statistical evidence of assocation between gender and career change influence.

2.  **Family influence**: the odds of career change interest for those had low family influence is 0.92 times the odds of those had no family influence (95%: 0.85 to 0.99, p=0.0269). No statistical differences comparing medium or high family influences to non family influence.

3.  **Current occupation**: the odds of career change interest for biologists is 1.19 times the odds for software developers (95%CI: 1.06 to 1.33, p=0.0026). Teachers and doctors also shows trends of higher odds of career change interests comparing to software developers.The other occupations do not have very different odds of career change interest as software developers.

4.  **Job security**: Higher Job security shows somewhat lower odds of career change interest (OR=0.99 for each unit change, 95%CI: 0.98 to 1, p=0.0926).

# Key Findings and Summary

## Findings:

-   \[Quick summary of findings from analysis #1\]

-   \[Quick summary of findings from analysis #2\]

-   \[Quick summary of findings from analysis #3\]

## Summary:

\[Summary of entire project\]
